---
title: "projectda"
author: "Mohit Parkhi"
date: "21 April 2018"
output:
  html_document:
    df_print: paged
---

```{r}
getwd()
eeg<-read.csv("C:/Users/mohit/Desktop/Data Analytics/eeg-brain-wave-for-confusion/EEG_data.csv")
 names(eeg)[14:15]=c("prelabel","userlabel")
summary(eeg)

print("Dimesions are:")
dim(eeg)


str(eeg)
sum(is.na(eeg$Attention))
```
Looking at the initial summary and structure, it is visible that all the all the features are numeric and there seem to be  no missing values. For this project, I want to find out what variables will have an effect on the confusion level. Also, I would like to predict if a student will get confused in the particular part of the video.
```{r}
#First I would create a box plot to find out the distribution of the status of the confusion levels as marked by the students.
library(ggplot2)

ggplot(data = eeg, aes(x = userlabel),fill=YesornO) +
  geom_bar(color = 'black',fill = I('red'))


# ggplot(eeg, aes( x = 1, y = userlabel ) ) +
               # geom_jitter(alpha = 0.1 ) +
               # geom_boxplot(alpha = 0.2, color = 'red' )

```
This shows that the data is uniformly distributed, there are equal number of times confusion have been felt and not.

Now for predicting whether a clip will be confuisng or not, we find out the correlation as to which feature will have maximum effect on the confusion level.
 Initially if we look at the data set we have to determine which feature will affect the confusion level the most. We can initially draw a corelation plot to see if the features are related to each other and also what is the correlation with the suer defined and predefined label with the features.
 
 

```{r}
library(corrplot)
corrplot(cor(eeg),method="number")

```
The above graph shows there is positve correaltion among the features, such as Delta, Theta, Alpha1, Alpha2, Beta1, Beta2, Gamma1, Gamma2. Also the features such as Delta, Theta, Aplha1, Alpha2, Beta1, Beta2 show positive correlation with the user defined confusion level. The subject Id and features like Delta., Theta, Alpha 1, Alpha2, Beta 1, Beta2, Gamma1, Gamma2 have positive correlation. then, there is negative correlation between Attention. mediation with the features and the levels Delta, Theta, Alpha 1, Alpha 2, Beta 1, Beta2, Gamma 1, Gamma2. Also, there is negative correlation with the feature Attention.



```{r}
#Trying the new correlation plot with the new data set after removing the outliers
# ------------------------------
# Load the data however you want
# ------------------------------
 
# I called my data dsBase, here I copy the data to dsBase.iqr
# I wanted to keep a copy of the original data set
eeg.iqr <- eeg
 
# Create a variable/vector/collection of the column names you want to remove outliers on.
vars <- c("Alpha1", "Alpha2", "Beta1", "Beta2","Gamma1","Gamma2")
 
# Create a variable to store the row id's to be removed
Outliers <- c()
 
# Loop through the list of columns you specified
for(i in vars){
 
  # Get the Min/Max values
  max <- quantile(eeg.iqr[,i],0.75, na.rm=TRUE) + (IQR(eeg.iqr[,i], na.rm=TRUE) * 1.5 )
  min <- quantile(eeg.iqr[,i],0.25, na.rm=TRUE) - (IQR(eeg.iqr[,i], na.rm=TRUE) * 1.5 )
  
  # Get the id's using which
  idx <- which(eeg.iqr[,i] < min | eeg.iqr[,i] > max)
  
  # Output the number of outliers in each variable
  print(paste(i, length(idx), sep=':'))
  
  # Append the outliers list
  Outliers <- c(Outliers, idx) 
}
 
# Sort, I think it's always good to do this
Outliers <- sort(Outliers)
 
# Remove the outliers
eeg.iqr <- eeg.iqr[-Outliers,]

summary(eeg.iqr)
str(eeg.iqr)
eeg.iqr$userlabel<-as.numeric(eeg.iqr$userlabel)

library(corrplot)
corrplot(cor(eeg.iqr),method="number")



```
The value changes for the correlation but there is no significant difference so we can conitnue with our original data too for analysis.

```{r}
#Now we can individually see how each feature affects the userlabel
eeg$userlabel<-as.numeric(eeg$userlabel)
eeg$userlabel<-factor(eeg$userlabel, ordered=T)
ggplot(data = eeg, aes(x = userlabel, y = Attention)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```


```{r}
ggplot(data = eeg, aes(x = userlabel, y = Delta)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```


```{r}
ggplot(data = eeg, aes(x = userlabel, y = Theta)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```


```{r}
sum<-summary(eeg)
typeof(sum)
range=((44780-6838)*1.5)+44780
typeof(range)
eeg_new <- subset(eeg, Alpha1 < (((44780-6838)*1.5)+44780))
#After removing the outliers
ggplot(data = eeg_new, aes(x = userlabel, y = Alpha1)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```







```{r}
library(gridExtra)
library(lattice)
eeg$userlabel<-factor(eeg$userlabel, ordered=T)
eeg.iqr$userlabel<-factor(eeg.iqr$userlabel, ordered=T)
grid.arrange(ggplot(data = eeg, aes(x = userlabel, y = Alpha2)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4),

ggplot(data = eeg.iqr, aes(x = userlabel, y = Alpha2)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
)

```




```{r}
ggplot(data = eeg, aes(x = userlabel, y = Beta1)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```



```{r}
ggplot(data = eeg, aes(x = userlabel, y = Beta2)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```





```{r}
ggplot(data = eeg, aes(x = userlabel, y = Gamma1)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```




```{r}
ggplot(data = eeg, aes(x = userlabel, y = Gamma2)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha=0.5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```
After looking at the correlation plots we plotted box plots for finding out the relation between each  feature with the user label. i.e confusion level with each feature, it can be observed that all the variables affect the user label wherein Attention, Delta, and theta values affect it the most. All other features affect but there is slight variation. Also, data is not distributed over all of the values of that feature. Moreover, all the features like Alpha, Beta, and Gamma affect Theta, and Delta and therefore  if Delta and Theta are considered they will be considered.
```{r}

#Random Forest
library(randomForest)
library(caret)

training_data1<-createDataPartition(eeg$userlabel,p=0.7,list=FALSE)
training_data<-eeg[training_data1,]
test_data<-eeg[-training_data1,]
training_data$userlabel=as.factor(training_data$userlabel)


rfmodel<-randomForest(userlabel~. , data=training_data)

rfmodel



```
 
```{r}

acc<-(3796+4361)/nrow(training_data)
acc
```
The accuracy on the model data comes out to be 90% which shows the model behaves on the model
```{r}
#predict 
test_data$userlabel<-as.factor(test_data$userlabel)

pred<-predict(rfmodel,newdata=test_data)

table(pred,test_data$userlabel)
```

```{r}
#calculating accuracy of test data
acc1<-(1655+1887)/nrow(test_data)
acc1


```

```{r}
#Selecting out important features
plot(rfmodel)

varImpPlot(rfmodel,
           sort = T,
           main="Variable Importance",
           n.var=9)
```

```{r}
#BAsed on the important variables according to random forest


rfmodel<-randomForest(userlabel~VideoID+SubjectID+Delta+Theta+Attention+Gamma1+Gamma2+Beta1+Beta2, data=training_data)

test_data$userlabel<-as.factor(test_data$userlabel)

pred<-predict(rfmodel,newdata=test_data)

table(pred,test_data$userlabel)
```

```{r}
acc_v<-(1728+1915)/nrow(test_data)
acc_v

#acc_d<-(1720+1896)/nrow(test_data)
#acc_d

#acc1<-(1703+1911)/nrow(test_data)
#acc1

confusionMatrix(data=test_data$userlabel,
                reference=test_data$userlabel)
```
The accuracy after taking important features out of all the features increases the accuracy from 92% to 94%
```{r}

#Since the output variable has binary output we use logistic regression

#Linear Modelling
library(memisc)
library(caret)
set.seed(1221)

training_data2<-createDataPartition(eeg$userlabel,p=0.7,list=FALSE)
training_data3<-eeg[training_data2,]
test_data1<-eeg[-training_data2,]

m1 <- glm(as.numeric(userlabel) ~ Theta, data = training_data3)
m2 <- update(m1, ~ . + Delta)
m3 <- update(m2, ~ . + Attention)


mtable(m1,m2,m3)


```




It was considered earlier that Theta affects the confusion level.
Theta, Delta, and Attention were considered to affect the confusion levels.
Other factors were correlated but affected Theta which in turn affected the confusion level.
SubjectID and VideoID came out to be important too from the feature importance of Random Forest which was not considered after correlation matrix was drawn.
